---
title: "p8105_hw2_cw3370"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(readxl)
library(lubridate)
library(knitr)
```

# Problem 1
### Read and clean the Mr.Trash Wheel sheet
```{r}
##specify the Mr. Trash Wheel sheet in the Excel file and to omit non-data entries
mr_trash_wheel_df <- 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "Mr. Trash Wheel",range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  ##omit rows that do not include dumpster-specific data
  drop_na(dumpster) %>% 
  ##round the number of sports balls to the nearest integer
  mutate(sports_balls = round(sports_balls))
```
<font size = 4 face="Times New Roman">Mr.Trash Wheel sheet includes `r nrow(mr_trash_wheel_df)` observations for `r ncol(mr_trash_wheel_df)` variables. For variables, they are `r colnames(mr_trash_wheel_df)`. 

`r colnames(mr_trash_wheel_df)[1]` is the number of dumpster. `r colnames(mr_trash_wheel_df)[2:4]` is the date of dumpster to be collected. `r colnames(mr_trash_wheel_df)[5:14]` describe the trash in the dumpster. 

The median number of sports balls in a dumpster in 2019 is `r median(pull(filter(mr_trash_wheel_df,year==2019),var = -2))`</font>


### Read and clean precipitation data for 2018 and 2019
```{r}
## read the data
## filter() to omit rows without precipitation data
## mutate() to add the variable year,change the variable month to character
Precipitation2018_df <- 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "2018 Precipitation",range = "A2:B14") %>% 
  janitor::clean_names() %>%
  mutate(month = month.name,year = 2018) %>% 
  drop_na(total)

Precipitation2019_df <- 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "2019 Precipitation",range = "A2:B14") %>% 
  janitor::clean_names() %>%
  mutate(month = month.name,year = 2019) %>% 
  drop_na(total)

##bind data together
Precipitation_bind_df <- 
  bind_rows(Precipitation2018_df,Precipitation2019_df) %>% 
  relocate(year)

```
<font size = 4 face="Times New Roman">For the precipitation data for 2018,2019, we collected `r nrow(Precipitation_bind_df)` month. 

The total precipitation in 2018 is `r sum(pull(filter(Precipitation_bind_df,year==2018),var = -1))`, including `r nrow(filter(Precipitation_bind_df,year==2018))` month. 

The total precipitation in 2019 is `r sum(pull(filter(Precipitation_bind_df,year==2019),var = -1))`, only including `r  nrow(filter(Precipitation_bind_df,year==2019))` month.</font>

# Problem 2
```{r}
#variable president:1 means the president was republican and -1 means the president was democratic
pols_month_df <- 
  read_csv("data/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon,into = c("year","month","day"),sep = "-") %>% 
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day),
         month = recode(month,`1` = month.name[1],`2` = month.name[2],`3` = month.name[3],`4` = month.name[4],`5` = month.name[5],`6` = month.name[6],`7` = month.name[7],`8` = month.name[8],`9` = month.name[9],`10` = month.name[10],`11` = month.name[11],`12` = month.name[12]),
         president = prez_gop - prez_dem) %>% 
  select(-prez_dem,-prez_gop,-day)
```
<font size = 4 face="Times New Roman">pols_month_df contains the data from 'pols_month.csv', including `r nrow(pols_month_df)` observations of `r ncol(pols_month_df)` related to the number of national politicians who are democratic or republican from  year `r min(select(pols_month_df,year))` to year `r `max(select(pols_month_df,year))`.

`r colnames(pols_month_df)[1:2]` are the month and day of count. `r colnames(pols_month_df)[3:5]` are the numbers of republican governors senator and representatives on the count date. `r colnames(pols_month_df)[6:8]` are the numbers of democratic governors senator and representatives on the count date. `r colnames(pols_month_df)[9]` are the indicators of the president, 1 means republican,-1 means democratic.</font>
```{r}
snp_df <- 
  read_csv("data/snp.csv") %>% 
  janitor::clean_names() %>%
  separate(date,into = c("month","day","year"),sep = "/") %>% 
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day)) %>% 
  select(-day) %>% 
  arrange(year,month) %>% 
  mutate(month = recode(month,`1` = month.name[1],`2` = month.name[2],`3` = month.name[3],`4` = month.name[4],`5` = month.name[5],`6` = month.name[6],`7` = month.name[7],`8` = month.name[8],`9` = month.name[9],`10` = month.name[10],`11` = month.name[11],`12` = month.name[12]),
         ) %>% 
  relocate(year,month)
```
<font size = 4 face="Times New Roman">snp_df contains the data from 'snp.csv', including `r nrow(snp_df)` observations of `r ncol(snp_df)` variables related to Standard & Poor’s stock market index (S&P) from year `r min(select(snp_df,year))` to year `r max(select(snp_df,year))`.

`r colnames(snp_df)[1:2]` are the year of month. `r colnames(snp_df)[3]`are the S&P.</font>
```{r}
unemployment_df <- 
  read_csv("data/unemployment.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment"
  ) %>% 
  mutate(month = recode(month,"jan" = month.name[1],"feb" = month.name[2],"mar" = month.name[3],"aapr" = month.name[4],"may" = month.name[5],"jun" = month.name[6],"jul" = month.name[7],"aug" = month.name[8],"sep" = month.name[9],"oct" = month.name[10],"nov" = month.name[11],"dec" = month.name[12]),
         year = as.integer(year))
```
<font size = 4 face="Times New Roman">unemployment_df contains the data from unemployment.csv, including `r nrow(unemployment_df)` observations of `r ncol(unemployment_df)` variables related to the percentage of unemployment from year `r min(select(unemployment_df,year))` to year `r max(select(unemployment_df,year))`.

`r colnames(unemployment_df)[1:2]` are the year and month. `r colnames(unemployment_df)[3]`are the percentage of unemployment.</font>
```{r}
## join by year and month
join_df <- 
  left_join(pols_month_df,snp_df,by = c("year","month")) %>% 
  left_join(unemployment_df,by =  c("year","month")) 

## use date to join
pols_month_df1 <- 
  mutate(
    pols_month_df,date = ymd(paste(as.character(year),month,1))
  )

snp_df1 <- 
  mutate(
    snp_df,date = ymd(paste(as.character(year),month,1))
  )

unemployment_df1 <- 
  mutate(
    unemployment_df,date = ymd(paste(as.character(year),month,1))
  )

join1_df <- 
  left_join(pols_month_df1,snp_df1,by = "date") %>% 
  left_join(unemployment_df1,by = "date") %>% 
  select(-year.y,-month.y,-year,-month,-date)

skimr::skim(join_df)
```

<font size = 4 face="Times New Roman">We use two ways to join these data. The first one is to left join by year and month. The second one is to create a Date variable by ymd() and join by this Date variable. The outcome is same.

join_df includes `r nrow(join_df)` observations of  `r ncol(join_df)` variables. It includes the number of national politicians who are democratic or republican from  year `r  min(select(pols_month_df,year))` to year`r max(select(pols_month_df,year))`, Standard & Poor’s stock market index (S&P) from year `r  min(select(snp_df,year))` to year `r max(select(snp_df,year))`,percentage of unemployment from year`r min(select(unemployment_df,year))` to year `r max(select(unemployment_df,year))`. 

The key variable in join_df is the same as three dataframe above. So l don't explain again.</font>

# Problem 3

<font size = 4 face="Times New Roman">Load and tidy the data.

Adjust the ethnicity variable and name_of_child variable, omit the repeated observations</font>
```{r}
popular_babyname_df1 <- 
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>%
  distinct() %>% 
  mutate(
    gender = str_to_title(gender),
    ethnicity = str_to_title(ethnicity), 
    ethnicity = recode(
      ethnicity,
      "Asian And Pacific Islander" = "Asian And Paci",
      "Black Non Hispanic" = "Black Non Hisp",
      "White Non Hispanic" = "White Non Hisp"
      ),
    childs_first_name = str_to_title(childs_first_name)
    )  

```

<font size = 4 face="Times New Roman">Popularity of the name "Oliva" as a female baby and popular male baby name</font>
```{r}
oliva_df <- 
  filter(popular_babyname_df1,childs_first_name == "Olivia",gender == "Female") %>% 
  select(-count,-gender,-childs_first_name) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    names_sort = TRUE,
    values_from = "rank"
  ) 

knitr::kable(oliva_df)

popular_male_name_df <- 
  filter(popular_babyname_df1,rank == 1,gender == "Male") %>% 
  select(-rank,-count,-gender) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    names_sort = TRUE,
    values_from = "childs_first_name"
  )

knitr::kable(popular_male_name_df)
```

<font size = 4 face="Times New Roman">Produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis) for male, white non-hispanic children born in 2016</font>
```{r}
male_wnh_2016_df <- 
  filter(popular_babyname_df1,ethnicity == "White Non Hisp",year_of_birth == 2016,gender == "Male") %>% 
  select(-ethnicity,-year_of_birth,-gender)

ggplot(data = male_wnh_2016_df, aes(x = rank ,y = count,size = count)) + 
    geom_point(size = 1) +
    labs(title = "Name ranks of male white non-hispanic children born in 2016" , x = "Rank of Name" , y = "Count of Name")

```


